---
title: "p8105_hw6_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "November 18, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
 
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw())
```

### Problem 1

#### 1-(1)
:Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. Modifiy victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r homicide_tidy_data}
homicide_data = 
  read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  janitor::clean_names() %>%  
  unite(city_state, city, state, sep = ",") %>%  
  mutate(hom_solved = ifelse((disposition == "Closed by arrest"), 1, 0)) %>% 
  filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"))  %>% 
  mutate(victim_race = 
           ifelse((victim_race == "White"), "white", "non-white"),
         victim_race = fct_relevel(victim_race, "white"),
         victim_age = as.numeric(victim_age)) 
```

#### 1-(2)
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r balti_logistic}
balti_logistic = 
  homicide_data %>% 
  filter(city_state == "Baltimore,MD") %>% 
  glm(hom_solved ~ victim_age + victim_sex + victim_race, 
      data = ., family = binomial())

balti_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(term = str_replace(term, "victim_racenon-white", "race:non-white")) %>%   knitr::kable(digits = 3)
```

#### 1-(3)
:Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of  purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r city_logistic}
city_homicide_data = 
  homicide_data %>% 
  group_by(city_state) %>% 
  nest()

city_hom_logistic =
  function(df) {
    glm(hom_solved ~ victim_age + victim_sex + victim_race, 
      family = binomial(), data = df)
  }

city_homicide_logistic_results =
  city_homicide_data %>%
  mutate(model = map(data, city_hom_logistic),
        tidy_results = model %>% map(broom::tidy)) %>%
  select(city_state, tidy_results) %>%
  unnest() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(city_state, term, OR, lower_CI_OR, upper_CI_OR) %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(term = str_replace(term, "victim_racenon-white", "race:non-white")) 

city_homicide_logistic_results 
```


#### 1-(4)
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r city_logistic_plot, fig.height = 13}
city_homicide_logistic_results %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  labs(
    title = "Odds ratio for solving homicides comparing non-white victims to white victims each city",
    subtitle = "Error bar using OR as center with 95% CIs",
    y = "Odds ratio",
    x = "City,State"
  ) +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 9)) +
  theme_bw()
```

In terms of solving homicides, non-white people has lower odds compared to white people in cities whose OR is less than 1. For cities whose OR is greater than 1, non-white people has greater odds compared to white people. However, 95% confidence interval contains 1, it is not statistically significant at the significance level 0.05. Most cities have no significant result from this reasons. Among cities whose CI does not contain 1, Boston,MA has the lowest odds ratio and Las Vegas,NV has the highest odds ratio but still less than 1.


### Problem 2

In this probelm, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

- babysex: baby’s sex (male = 1, female = 2)
- bhead: baby’s head circumference at birth (centimeters)
- blength: baby’s length at birth (centimeteres)
- bwt: baby’s birth weight (grams)
- delwt: mother’s weight at delivery (pounds)
- fincome: family monthly income (in hundreds, rounded)
- frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
- gaweeks: gestational age in weeks
- malform: presence of malformations that could affect weight (0 = absent, 1 = present)
- menarche: mother’s age at menarche (years)
- mheigth: mother’s height (inches)
- momage: mother’s age at delivery (years)
- mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
- parity: number of live births prior to this pregnancy
- pnumlbw: previous number of low birth weight babies
- pnumgsa: number of prior small for gestational age babies
- ppbmi: mother’s pre-pregnancy BMI
- ppwt: mother’s pre-pregnancy weight (pounds)
- smoken: average number of cigarettes smoked per day during pregnancy
- wtgain: mother’s weight gain during pregnancy (pounds)

#### 2-(1)
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r birthweight_data_load}
race_level = c("White", "Black", "Asian", "Puerto Rican", "Other")

birthweight_data = 
  read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = as.factor(babysex),
         frace = case_when(
             frace == 1 ~ "White",
             frace == 2 ~ "Black",
             frace == 3 ~ "Asian",
             frace == 4 ~ "Puerto Rican",
             frace == 8 ~ "Other"),
         frace = fct_relevel(frace, race_level),
         malform = as.factor(malform),
         mrace = case_when(
             mrace == 1 ~ "White",
             mrace == 2 ~ "Black",
             mrace == 3 ~ "Asian",
             mrace == 4 ~ "Puerto Rican",
             mrace == 8 ~ "Other"),
         mrace = fct_relevel(mrace, race_level))

# Check missing data
sum(is.na(birthweight_data))
```


#### 2-(2)
:Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use  add_predictions and add_residuals in making this plot.

**1) Association: bwt vs variables**
```{r fig.width = 3, fig.asp = .6, out.width = "50%"}

## Association: bwt vs variables

birthweight_data %>% 
  ggplot(aes(x = babysex, y = bwt)) +
  geom_boxplot() +
  labs(subtitle = "birthweight vs baby’s sex") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect 

birthweight_data %>% 
  ggplot(aes(x = bhead, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs baby’s head circumference at birth",
       y = "birthweight",
       x = "baby's head circumference (cm)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## strong 
  
birthweight_data %>% 
  ggplot(aes(x = blength, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs baby’s length at birth (cm)",
       y = "birthweight",
       x = "baby’s length at birth (cm)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## strong

birthweight_data %>% 
  ggplot(aes(x = delwt, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s weight at delivery",
       y = "birthweight",
       x = "mother’s weight at delivery (lb)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week

birthweight_data %>% 
  ggplot(aes(x = fincome, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs family monthly income",
       y = "birthweight",
       x = "family monthly income") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = frace, y = bwt)) +
  geom_boxplot() +
  labs(subtitle = "birthweight vs father's race",
       y = "birthweight",
       x = "father's race") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = gaweeks, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs gestational age in weeks",
       y = "birthweight",
       x = "gestational age in weeks") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week

birthweight_data %>% 
  ggplot(aes(x = malform, y = bwt)) +
  geom_boxplot() +
  labs(subtitle = "birthweight vs presence of malformations",
       y = "birthweight",
       x = "presence of malformations") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = menarche, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s age at menarche",
       y = "birthweight",
       x = "mother’s age at menarche (years)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = mheight, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s height",
       y = "birthweight",
       x = "mother’s height (inches)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = momage, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s age at delivery",
       y = "birthweight",
       x = "mother’s age at delivery (years)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = mrace, y = bwt)) +
  geom_boxplot() +
  labs(subtitle = "birthweight vs mother's race",
       y = "birthweight",
       x = "mother's race") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week effect

birthweight_data %>% 
  ggplot(aes(x = parity, y = bwt)) +
  geom_point() +
  labs(subtitle = "birthweight vs number of live births prior to this pregnancy",
       y = "birthweight",
       x = "number of live births prior to this pregnancy") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))  ## no effect

birthweight_data %>% 
  ggplot(aes(x = pnumlbw, y = bwt)) +
  geom_point() +
  labs(subtitle = "birthweight vs previous number of low birth weight babies",
       y = "birthweight",
       x = "previous number of low birth weight babies") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))  ## no effect

birthweight_data %>% 
  ggplot(aes(x = pnumsga, y = bwt)) +
  geom_point() +
  labs(subtitle = "birthweight vs number of prior small for gestational age babies",
       y = "birthweight",
       x = "number of prior small for gestational age babies") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))  ## no effect

birthweight_data %>% 
  ggplot(aes(x = ppbmi, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s pre-pregnancy BMI",
       y = "birthweight",
       x = "mother’s pre-pregnancy BMI") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week effect

birthweight_data %>% 
  ggplot(aes(x = ppwt, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s pre-pregnancy weight ",
       y = "birthweight",
       x = "mother’s pre-pregnancy weight (lb)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week effect

birthweight_data %>% 
  ggplot(aes(x = smoken, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs average number of cigarettes smoked per day during pregnancy",
       y = "birthweight",
       x = "average number of cigarettes smoked per day during pregnancy") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## no effect

birthweight_data %>% 
  ggplot(aes(x = wtgain, y = bwt)) +
  geom_point() + 
  geom_smooth() +
  labs(subtitle = "birthweight vs mother’s weight gain during pregnancy",
       y = "birthweight",
       x = "mother’s weight gain during pregnancy (lb)") +
  theme(plot.subtitle = element_text(size = 9),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7)) ## week effect
```

With regard to association between outcome(birthweight) and covariates, in case of categorical covariates, boxplots are used and in case of continuous covariates, I looked at the scatter plot with loess smoother. As a result, I categorized covariates with the level of association with birthweight:
- **strong association**: bhead, blength
- **week association**: delwt, gaweeks, mrace, ppbmi, ppwt, wtgain
- **no association**: babysex, fincome, frace, malform, menarche, mheight, momage, parity, pnumlbw, pnumbsga, smoken

 
**2) Model comparison: Hypothesis testing**
```{r}
## strong effect variables only
fit_str = 
  birthweight_data %>%
  select(bwt, bhead, blength) %>% 
  lm(bwt ~., data = .)

## strong + week effect variables 
fit_str_week =  
  birthweight_data %>% 
  select(bwt, bhead, blength, delwt, gaweeks, mrace, ppbmi, ppwt, wtgain) %>% 
  lm(bwt ~., data = .)
summary(fit_str_week)

anova(fit_str_week, fit_str) %>% 
  broom::tidy() %>% 
  knitr::kable() 
```

After skimming the association between birthweight and each explanatory variables, I fit two linear models as follows:
- 1) birthweight with variables categorized in strong association.
- 2) birthweight with variables categorized in strong and week association.
As a result of model comparison, the more complex model (model 2)) is significantly better than the simpler model (model 1)) at the significance level 0.05. 

**3) plot of model residuals against fitted values** 

```{r fig.width = 4, fig.asp = .6, out.width = "50%"}
birthweight_data %>% 
  add_predictions(fit_str_week) %>% 
  add_residuals(fit_str_week) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .25) +
  geom_smooth(se = FALSE) +
  labs(title = "Model residuals vs fitted values",
       x = "Predicted value",
       y = "Residuals")
```

This model is poor if birthweight is low since residual is much greater than 0 but model fits well if birthweight is over 2000 with residuals close 0. 

#### 2-(3)
Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these.

```{r warning = FALSE}
fit_alt_main =
  birthweight_data %>%
  select(bwt, blength, gaweeks) %>% 
  lm(bwt ~., data = .)

fit_alt_int =
  lm(bwt ~ bhead*blength + bhead*babysex + blength*babysex, 
     data = birthweight_data)

crossv_mc(birthweight_data, 100) %>% 
  mutate(str_week      = map(train, ~lm(bwt ~ bhead + blength + delwt + gaweeks + mrace + ppbmi + ppwt + wtgain, data = .x)),
         alt_main = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         alt_int  = map(train, ~lm(bwt ~ bhead*blength + bhead*babysex +
                                     blength*babysex, data = .x))) %>% 
  mutate(rmse_str_week = map2_dbl(str_week, test, ~rmse(model = .x, data = .y)),
         rmse_alt_main = map2_dbl(alt_main, test, ~rmse(model = .x, data = .y)),
         rmse_alt_int = map2_dbl(alt_int, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(title = "root mean squared errors (RMSEs) in each models")
```

Based on this result, final model (str_week) is improved in predictive accuracy than main effect only model and a model including three-way interaction. 
