---
title: "p8105_hw6_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "November 18, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw())
```

### Problem 1

#### 1-(1)
:Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. Modifiy victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r homicide_tidy_data}
homicide_data = 
  read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  janitor::clean_names() %>%  
  unite(city_state, city, state, sep = ",") %>%  
  mutate(hom_solved = ifelse((disposition == "Closed by arrest"), 1, 0)) %>% 
  filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"))  %>% 
  mutate(victim_race = 
           ifelse((victim_race == "White"), "white", "non-white"),
         victim_race = fct_relevel(victim_race, "white"),
         victim_age = as.numeric(victim_age)) 
```

#### 1-(2)
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r balti_logistic}
balti_logistic = 
  homicide_data %>% 
  filter(city_state == "Baltimore,MD") %>% 
  glm(hom_solved ~ victim_age + victim_sex + victim_race, 
      data = ., family = binomial())

balti_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(term = str_replace(term, "victim_racenon-white", "race:non-white")) %>%   knitr::kable()
```

#### 1-(3)
:Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of  purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r city_logistic}
city_homicide_data = 
  homicide_data %>% 
  group_by(city_state) %>% 
  nest()

city_hom_logistic =
  function(df) {
    glm(hom_solved ~ victim_age + victim_sex + victim_race, 
      family = binomial(), data = df)
  }

city_homicide_logistic_results =
  city_homicide_data %>%
  mutate(model = map(data, city_hom_logistic),
        tidy_results = model %>% map(broom::tidy)) %>%
  select(city_state, tidy_results) %>%
  unnest() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(city_state, term, OR, lower_CI_OR, upper_CI_OR) %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(term = str_replace(term, "victim_racenon-white", "race:non-white")) 

city_homicide_logistic_results 
```


#### 1-(4)
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r city_logistic_plot, fig.height = 7}
city_homicide_logistic_results %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  labs(
    title = "Odds ratio for solving homicides comparing non-white victims to white victims each city",
    subtitle = "Error bar using OR as center with 95% CIs",
    y = "Odds ratio",
    x = "City,State"
  ) +
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 9)) +
  theme_bw()
```

In terms of solving homicides, non-white people has lower odds compared to white people in cities whose OR is less than 1. For cities whose OR is greater than 1, non-white people has greater odds compared to white people. However, if 95% confidence interval contains 1, it is not statistically significant at the significance level 0.05. 


### Problem 2

In this probelm, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

- babysex: baby’s sex (male = 1, female = 2)
- bhead: baby’s head circumference at birth (centimeters)
- blength: baby’s length at birth (centimeteres)
- bwt: baby’s birth weight (grams)
- delwt: mother’s weight at delivery (pounds)
- fincome: family monthly income (in hundreds, rounded)
- frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
- gaweeks: gestational age in weeks
- malform: presence of malformations that could affect weight (0 = absent, 1 = present)
- menarche: mother’s age at menarche (years)
- mheigth: mother’s height (inches)
- momage: mother’s age at delivery (years)
- mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
- parity: number of live births prior to this pregnancy
- pnumlbw: previous number of low birth weight babies
- pnumgsa: number of prior small for gestational age babies
- ppbmi: mother’s pre-pregnancy BMI
- ppwt: mother’s pre-pregnancy weight (pounds)
- smoken: average number of cigarettes smoked per day during pregnancy
- wtgain: mother’s weight gain during pregnancy (pounds)

#### 2-(1)
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
#?????????????????????frace, mrace fct order??????
```{r birthweight_data_load}
birthweight_data = 
  read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         frace = case_when(
             frace == 1 ~ "White",
             frace == 2 ~ "Black",
             frace == 3 ~ "Asian",
             frace == 4 ~ "Puerto Rican",
             frace == 8 ~ "Other"),
         malform = as.factor(malform),
         mrace = as.factor(mrace))
```


#### 2-(2)
:Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use  add_predictions and add_residuals in making this plot.

**1) Association: bwt vs variables**
```{r fig.width = 3, fig.asp = .6, out.width = "50%"}

## Association: bwt vs variables

birthweight_data %>% 
  ggplot(aes(x = babysex, y = bwt)) +
  geom_point() ## week 

birthweight_data %>% 
  ggplot(aes(x = bhead, y = bwt)) +
  geom_point() + geom_smooth() ## strong 
  
birthweight_data %>% 
  ggplot(aes(x = blength, y = bwt)) +
  geom_point() + geom_smooth() ## strong

birthweight_data %>% 
  ggplot(aes(x = delwt, y = bwt)) +
  geom_point() + geom_smooth() ## week

birthweight_data %>% 
  ggplot(aes(x = fincome, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = frace, y = bwt)) +
  geom_point() ## week effect

birthweight_data %>% 
  ggplot(aes(x = gaweeks, y = bwt)) +
  geom_point() + geom_smooth() ## week

birthweight_data %>% 
  ggplot(aes(x = malform, y = bwt)) +
  geom_point() ## week effect

birthweight_data %>% 
  ggplot(aes(x = menarche, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = mheight, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = momage, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = mrace, y = bwt)) +
  geom_point() ## week effect

birthweight_data %>% 
  ggplot(aes(x = parity, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = pnumlbw, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = pnumsga, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = ppwt, y = bwt)) +
  geom_point() + geom_smooth() ## week effect

birthweight_data %>% 
  ggplot(aes(x = smoken, y = bwt)) +
  geom_point() + geom_smooth() ## no effect

birthweight_data %>% 
  ggplot(aes(x = wtgain, y = bwt)) +
  geom_point() + geom_smooth() ## week effect
```


**2) Model comparison: Hypothesis testing**
```{r}
## strong effect variables only
fit_null = 
  birthweight_data %>%
  select(bwt, bhead, blength) %>% 
  lm(bwt ~., data = .)

## strong + week effect variables 
fit_alt =  
  birthweight_data %>% 
  select(bwt, babysex, bhead, blength, delwt, frace, gaweeks, malform, mrace, ppbmi, ppwt, wtgain) %>% 
  lm(bwt ~., data = .)

anova(fit_null, fit_alt) %>% 
  broom::tidy() %>% 
  knitr::kable() 
```

First, I looked at the association between outcome variable(birthweight) and each explanatory variable by scatterplot. After skimming, I fit the linear model with variables which are considered strongly associated with outcome variable based on the scatterplots. And then, I added variables which has shown week association with outcome variable. As a result of hypothesis testing, the more complex model is significantly better than the simpler model at the significance level 0.05. 

**3) plot of model residuals against fitted values** 

```{r fig.width = 4, fig.asp = .6, out.width = "50%"}
birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  mutate(babysex = ifelse((babysex == 1), "male", "female")) %>% 
  ggplot(aes(x = babysex, y = resid)) + 
  geom_violin(aes(fill = babysex), alpha = .5) +
  labs(x = "baby sex",
       y = "residuals") +
  viridis::scale_fill_viridis(
    discrete = TRUE
    )

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = bhead, y = resid)) + 
  geom_point() +
  labs(x = "baby's head circumference at birth (cm)",
       y = "residuals") 

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = blength, y = resid)) + 
  geom_point() +
  labs(x = "baby's length at birth (cm)",
       y = "residuals") 

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = delwt, y = resid)) + 
  geom_point() +
  labs(x = "mother's weight at delivery(lb)",
       y = "residuals") 

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = frace, y = resid)) + 
  geom_violin(aes(fill = frace), alpha = .5) +
  labs(x = "father's race",
       y = "residuals") +
  viridis::scale_fill_viridis(
    discrete = TRUE
    )

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = gaweeks, y = resid)) + 
  geom_point() +
  labs(x = "gestational ages in weeks",
       y = "residuals")

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  mutate(malform = ifelse((malform == 0), "absent", "present")) %>% 
  ggplot(aes(x = malform, y = resid)) + 
  geom_violin(aes(fill = malform), alpha = .5) +
  labs(x = "presence of malformations",
       y = "residuals") +
  viridis::scale_fill_viridis(
    discrete = TRUE
    )

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = mrace, y = resid)) + 
  geom_violin(aes(fill = mrace), alpha = .5) +
  labs(x = "mather's race",
       y = "residuals") +
  viridis::scale_fill_viridis(
    discrete = TRUE
    )

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = ppbmi, y = resid)) + 
  geom_point() +
  labs(x = "mother's pre-pregnancy BMI",
       y = "residuals")

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = ppwt, y = resid)) + 
  geom_point() +
  labs(x = "mother's pre-pregnancy weight(lb)",
       y = "residuals")

birthweight_data %>% 
  add_residuals(fit_alt) %>% 
  ggplot(aes(x = wtgain, y = resid)) + 
  geom_point() +
  labs(x = "mother's weight gain during pregnancy",
       y = "residuals")
```


#### 2-(3)????????????????????????????????
Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these.
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in  purrr as appropriate. 

```{r}
fit_alt_main =
  birthweight_data %>%
  select(bwt, blength, gaweeks) %>% 
  lm(bwt ~., data = .)

fit_alt_int =
  lm(bwt ~ bhead*blength + bhead*babysex + blength*babysex, 
     data = birthweight_data)

crossv_mc(birthweight_data, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble)) %>% 
  mutate(alt      = map(train, 
                   ~lm(bwt ~ babysex + bhead + blength + delwt + frace + gaweeks
                       + malform + mrace + ppbmi + ppwt + wtgain, data = .x)),
         alt_main = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         alt_int  = map(train, ~lm(bwt ~ bhead*blength + bhead*babysex +
                                     blength*babysex, data = .x))) %>% 
  mutate(rmse_alt    = map2_dbl(alt, test, ~rmse(model = .x, data = .y)),
         rmse_alt_main = map2_dbl(alt_main, test, ~rmse(model = .x, data = .y)),
         rmse_alt_int = map2_dbl(alt_int, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

```


